---
title: "assignment_2"
author: "JJayes"
date: "26/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(scales)
library(rvest)
library(readxl)
library(here)

theme_set(theme_light())
```


## Agricultural productivity

```{r}

df <- read_excel(here("data", "Assignment_2", "agric_tfp.xlsx")) %>% janitor::clean_names()

df <- df %>% 
  pivot_longer(-c(order, fao, iso3, country_territory, region, sub_region, inc_i), names_to = "year") %>% 
  mutate(year = parse_number(year))

```



```{r}
df %>%
  ggplot(aes(year, value, colour = country_territory)) +
  geom_point() +
  geom_line() +
  theme(legend.position = "none") +
  gghighlight::gghighlight(region == "SSA")
```

### Change base year

```{r}
df %>% 
  filter(country_territory == "Ghana") %>% 
  ggplot(aes(year, value)) +
  geom_line() +
  geom_point()
```

```{r}
df_1961 <- df %>% 
  filter(year == 1961) %>% 
  select(iso3, value_1961 = value)

df <- df %>% 
  left_join(df_1961, by = c("iso3"))

df <- df %>% 
  mutate(value_base_1961 = value / value_1961 * 100)
```


Test

```{r}
df %>% 
  filter(country_territory == "Ghana") %>% 
  ggplot(aes(year, value_base_1961)) +
  geom_line() +
  geom_point()
```

```{r}
df %>% 
  # filter(country_territory == "Ghana") %>% 
  ggplot(aes(year, value_base_1961, colour = country_territory)) +
  geom_line() +
  geom_point() +
  facet_wrap(~ region) +
  theme(legend.position = "none")
```


## Shares of GDP

```{r}
df <- read_excel(here("data", "Assignment_2", "sectoral_shares_of_gdp.xlsx")) %>% janitor::clean_names()

df <- df %>% 
  pivot_longer(-c(country_name, country_code, indicator_name, indicator_code), names_to = "year") %>% 
  mutate(year = parse_number(year))

```

```{r}
# df %>% 
#   filter(country_name == "South Africa") %>% 
#   ggplot(aes(year, value, fill = indicator_name)) +
#   geom_area(position = "fill")
```

### Creating shre of agric var

```{r}
df_agric <- df %>% 
  select(country_name, indicator_name, year, value) %>% 
  pivot_wider(names_from = indicator_name) %>% 
  janitor::clean_names() %>% 
  rowwise() %>% 
  mutate(sum_shares = sum(agriculture_forestry_and_fishing_value_added_percent_of_gdp, 
                          manufacturing_value_added_percent_of_gdp, 
                          services_value_added_percent_of_gdp,
                          industry_including_construction_value_added_percent_of_gdp,
                          na.rm = T)) %>% 
  mutate(share_agric = agriculture_forestry_and_fishing_value_added_percent_of_gdp / sum_shares*100) %>% 
  select(country_name, year, share_agric)
```

```{r}
df <- df %>% 
  inner_join(df_agric, by = c("country_name", "year"))
```

### Need a list of countries with region 

```{r}
library(countrycode)

df <- df %>% 
  mutate(continent = countrycode(country_code, "iso3c", "continent"))
```


```{r}
df %>% 
  filter(year == 1990,
         !is.na(value)) %>% 
  mutate(country_name = fct_reorder(country_name, share_agric)) %>% 
  ggplot(aes(value, country_name, fill = indicator_name)) +
  geom_col(position = "fill")

countries <- df %>% 
  filter(
    year %in% c(1990, 2020),
    continent %in% c("Americas", "Africa", "Asia"),
    indicator_name == "Agriculture, forestry, and fishing, value added (% of GDP)"
  ) %>%
  select(country_name, year, value, continent) %>% 
  pivot_wider(names_from = year, values_from = value) %>% 
  filter(!is.na(`1990`),
         !is.na(`2020`)) %>% 
  select(continent, country_name) %>% 
  distinct() %>% 
  group_by(continent) %>% 
  sample_n(size = 15) %>% 
  ungroup() %>% 
  select(country_name) %>% pull()

jpeg(filename = here("images", "assignment_2", "gdp_comp_199_2020.jpeg"),
     height = 8,
     width = 8,
     units = "in",
     res = 1000)

df %>%
  filter(
    year %in% c(1990, 2020),
    # continent %in% c("Americas", "Africa", "Asia"),
    country_name %in% countries
  ) %>%
  mutate(indicator_name = str_remove(indicator_name, ", value added \\(\\% of GDP\\)")) %>% 
  mutate(country_name = fct_reorder(country_name, share_agric)) %>%
  ggplot(aes(value, country_name, fill = indicator_name)) +
  geom_col(position = "fill") +
  facet_grid(continent ~ year, scales = "free") +
  scale_fill_brewer(palette = "Spectral", direction = -1) +
  scale_x_continuous(labels = percent_format()) +
  labs(x = "Share of GDP (value added) from sector",
       y = NULL,
       fill = NULL) +
  theme(legend.position = "bottom")

dev.off()
```

### Shares of employment

```{r}
df <- read_excel(here("data", "Assignment_2", "sectoral_shares_of_employment.xlsx")) %>% janitor::clean_names()

df <- df %>% 
  pivot_longer(-c(country_name, country_code, indicator_name, indicator_code), names_to = "year") %>% 
  mutate(year = parse_number(year))

df <- df %>% 
  mutate(continent = countrycode(country_code, "iso3c", "continent"))

df_agric <- df %>% 
  select(country_name, indicator_name, year, value) %>% 
  pivot_wider(names_from = indicator_name) %>% 
  janitor::clean_names() %>%
  rowwise() %>% 
  mutate(sum_shares = sum(employment_in_agriculture_percent_of_total_employment_modeled_ilo_estimate,
                          employment_in_industry_percent_of_total_employment_modeled_ilo_estimate,
                          employment_in_services_percent_of_total_employment_modeled_ilo_estimate,
                          na.rm = T)) %>% 
  mutate(share_agric = employment_in_agriculture_percent_of_total_employment_modeled_ilo_estimate / sum_shares*100) %>% 
  select(country_name, year, share_agric)

df <- df %>% 
  inner_join(df_agric, by = c("country_name", "year"))

```

```{r}
# df %>% 
#   filter(country_name == "India") %>% 
#   ggplot(aes(year, value, fill = indicator_name)) +
#   geom_col()
```

```{r}

countries <- df %>% 
  filter(continent %in% c("Americas", "Africa", "Asia"),
          year %in% c(1991, 2018),
         !is.na(value)) %>% 
  select(continent, country_name) %>% 
  distinct() %>% 
  group_by(continent) %>% 
  sample_n(size = 15) %>% 
  ungroup() %>% 
  select(country_name) %>% pull()

jpeg(filename = here("images", "assignment_2", "employment_1991_2018.jpeg"),
     height = 8,
     width = 8,
     units = "in",
     res = 1000)

df %>%
  filter(
    year %in% c(1991, 2018),
    continent %in% c("Americas", "Africa", "Asia")
  ) %>%
  filter(!is.na(value),
         country_name %in% countries) %>%
  mutate(country_name = fct_reorder(country_name, share_agric),
         indicator_name = str_remove(indicator_name, "\\(.*")) %>%
  ggplot(aes(value, country_name, fill = indicator_name)) +
  geom_col(position = "fill") +
  scale_fill_brewer(palette = "Dark2") +
  facet_grid(continent ~ year, scales = "free") +
  scale_x_continuous(labels = percent_format()) +
  theme(legend.position = "bottom") +
  labs(x = NULL,
       y = NULL,
       fill = NULL)

dev.off()
```

```{r}
df_agric_to_join <- df %>% 
  mutate(value = share_agric) %>% 
  filter(str_detect(indicator_name, "agricul")) %>% 
    select(-continent, -share_agric)
```


## Gapminder style chart with GDP per capita and 

```{r}
df <- read_excel(here("data", "Assignment_2", "gdp_per_capita.xlsx")) %>% janitor::clean_names()

df <- df %>% 
  pivot_longer(-c(country_name, country_code, indicator_name, indicator_code), names_to = "year") %>% 
  mutate(year = parse_number(year))
```

### population to join

```{r}
# df <- read_excel(here("data", "Assignment_2", "population.xlsx")) %>% janitor::clean_names()
# 
# df <- df %>% 
#   pivot_longer(-c(country_name, country_code, indicator_name, indicator_code), names_to = "year") %>% 
#   mutate(year = parse_number(year))
```


```{r}
df <- df %>% 
  bind_rows(df_agric_to_join) %>% 
  filter(year == 2015) %>% 
  select(country_name, indicator_name, value) %>% 
  mutate(continent = countrycode(country_name, "country.name", "continent")) %>% 
  filter(!is.na(continent)) %>% 
  pivot_wider(names_from = indicator_name, values_from = value) %>% 
  janitor::clean_names()


jpeg(filename = here("images", "assignment_2", "agric_gdp_2015.jpeg"),
     height = 5,
     width = 8,
     units = "in",
     res = 1000)

df %>% 
  ggplot(aes(employment_in_agriculture_percent_of_total_employment_modeled_ilo_estimate, 
             gdp_per_capita_current_us)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(aes(colour = continent, shape = continent)) +
  scale_y_log10(labels = dollar_format()) +
  scale_x_continuous(labels = percent_format(scale = 1)) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Percentage of workforce employed in Agriculture",
       y = "GDP per capita (2011 US$)(log scale)",
       colour = NULL,
       shape = NULL)

dev.off()
```

### Evolution over time

```{r}
df_agric_to_join %>% 
  filter(country_name %in% c("Vietnam", "Ethiopia")) %>% 
  ggplot(aes(year, value, colour = country_name)) +
  geom_line() +
  geom_point()

jpeg(filename = here("images", "assignment_2", "agric_vs_gdp__pc.jpeg"),
     height = 5,
     width = 8,
     units = "in",
     res = 1000)

df_agric_to_join %>% 
  inner_join(df %>% 
  filter(country_name %in% c("Vietnam", "Ethiopia")) %>% 
  select(year, country_name, gdp_pc = value), by = c("year", "country_name")) %>% 
  ggplot(aes(value, gdp_pc, colour = country_name, label = year)) +
  geom_line() +
  geom_point(size = 2) +
  geom_text(check_overlap = T, show.legend = F, hjust = 1.5) +
  scale_x_continuous(labels = percent_format(scale = 1)) +
  scale_y_continuous(labels = dollar_format()) +
  scale_color_brewer(palette = "Dark2") +
  expand_limits(x = 35) +
  labs(x = "Share of labour force in agriculture",
       y = "GDP per capita (1990$)",
       colour = "Country")

dev.off()
```


## Another fun one

Could do a one with share of GDP in agriculture and share of employment in agriculture and share agric productivity.


```{r}



```


### Urbanization

```{r}

df <- read_excel(here("data", "Assignment_2", "urbanization.xlsx")) %>% janitor::clean_names()

df <- df %>% 
  pivot_longer(-c(country_name, country_code, indicator_name, indicator_code), names_to = "year") %>% 
  mutate(year = parse_number(year))

df <- df %>% 
  mutate(continent = countrycode(country_code, "iso3c", "continent"))

```


```{r}
jpeg(filename = here("images", "assignment_2", "urbanization.jpeg"),
     height = 5,
     width = 8,
     units = "in",
     res = 1000)

df %>% 
  filter(value != 100,
         continent %in% c("Americas", "Africa", "Asia")) %>% 
  ggplot(aes(year, value, colour = country_name)) +
  # geom_point() +
  geom_line(size = 1.2) +
  facet_wrap(~ continent) +
  gghighlight::gghighlight(country_name %in% 
                             c("South Africa", "Brazil", "India", "China"), 
                           calculate_per_facet = T) +
  scale_x_continuous() +
  scale_y_continuous(labels = percent_format(scale = 1)) +
  labs(x = NULL,
       y = "Share of population living in urban area")

dev.off()
```

