---
title: "ingest"
author: "JJayes"
date: "14/11/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(scales)
library(here)

theme_set(theme_light())

source(here("code", "fns", "plot_variable.R"))


df <- read_rds(here("data", "df.rds"))

```

## Purpose

I found this dataset - Historical Cross-Country Technology Adoption (HCCTA) Dataset online. It's on the NBER website. From Diego A. Comin and Bart Hobijin.

Here's the [link](https://www.nber.org/research/data/historical-cross-country-technology-adoption-hccta-dataset)

The data description [here](https://data.nber.org/hccta/hcctadhelp.pdf)

### Ingest and clean

```{r}
df <- read.csv(here("data", "hcctad.txt")) %>% as_tibble() %>% janitor::clean_names()

df <- df %>% 
  mutate(across(.cols = -c(variable, year), .fns = parse_number))

df %>% distinct(variable)
```

### Reshape

```{r}
df <- df %>% 
  pivot_longer(-c(variable, year), names_to = "country", values_to = "value") %>% 
  mutate(country = str_to_title(str_replace_all(country, "_", " ")))
```

### Cleaning

```{r}
# df %>% 
#   filter(country == "Germany",
#          variable == "Length of Railway line open") %>% 
#   ggplot(aes(year, value)) +
#   geom_line()

df <- df %>% 
  mutate(value = case_when(
    country == "Germany" & variable == "Length of Railway line open" & value > 4e06 ~ 0,
    TRUE ~ value))
```

### Skim

```{r}
df_skim <- df %>% pivot_wider(names_from = "variable") %>% skimr::skim()
```

### What variables are complete?

```{r}
df_skim %>% select(skim_variable, complete_rate) %>% 
  mutate(skim_variable = fct_reorder(skim_variable, complete_rate)) %>% 
  head(20) %>% 
  ggplot(aes(complete_rate, skim_variable, fill = complete_rate)) +
  geom_col() +
  scale_x_continuous(labels = percent_format())
```

What kinds of variables are there?

Standard with high complete rate:

- year
- pop
- real and nom. GDP

Some interesting continuous variables

- Length of Railway line open
- Freight traffic on railways
- Passenger traffic on railways

- Energy output - how is this measured?

Types of technology

- Phones

### Modify steel

What are the different compositions of steel production by country

```{r}
df %>% 
  filter(str_detect(variable, "steel"),
         variable != "steel total",
         !is.na(value)) %>% 
  mutate(country = fct_lump(country, 9)) %>% 
  filter(!country == "Other") %>% 
  pivot_wider(names_from = variable, values_fill = 0) %>% 
  pivot_longer(-c(year, country), names_to = "variable") %>% 
  mutate(country = fct_reorder(country, value, .desc = T, .fun = sum),
         variable = str_to_title(str_remove(variable, "steel - "))) %>% 
  ggplot(aes(year, value, fill = variable, group = variable)) +
  geom_area() +
  scale_y_continuous(labels = number_format()) +
  facet_wrap(~ country, scales = "free") +
  labs(x = NULL,
       y = "Thousands of metric tons of crude steel",
       fill = "Steel production method")
```

So the issue is that it's great to have all these types of steel over time. but wouldn't it be nicer just to have a sum.

```{r}
df %>% filter(str_detect(variable, "steel")) %>%
  group_by(year, country) %>% 
  mutate(steel_total = sum(value, na.rm = T)) %>% 
  ungroup() %>% 
  filter(steel_total > 0) %>% 
  ggplot(aes(year, steel_total, colour = country)) +
  geom_point() +
  geom_line()

df_steel <- df %>% filter(str_detect(variable, "steel")) %>%
  group_by(year, country) %>% 
  mutate(steel_total = sum(value, na.rm = T)) %>% 
  ungroup() %>% 
  select(year, country, value = steel_total) %>% 
  mutate(variable = "steel total") %>% 
  filter(value > 0)

df <- df %>% 
  bind_rows(df_steel)

# df %>% write_rds(here("data", "df.rds"))
```

What does steel total look like??

```{r}
df %>% 
  filter(variable == "steel total") %>% 
  ggplot(aes(year, value, colour = country)) +
  geom_point() +
  geom_line()
```

What about by decade

```{r}
df %>% 
  filter(variable == "steel total") %>% 
  mutate(decade = year - year %% 10) %>% 
  group_by(country, decade) %>% 
  mutate(value = sum(value, na.rm = T)) %>% 
  ggplot(aes(decade, value, fill = country)) +
  geom_col(position = "fill") 
```


```{r}
# df <- read_rds(here("data", "df.rds"))
```


### Data descriptions

```{r}
# library(readxl)
# hccta_data_description <- read_excel("data/hccta_data_description.xlsx")
# 
# df_data_descs <- hccta_data_description %>% 
#   separate(key, into = c("variable", "description"), sep = ", ", extra = "merge")
# 
# df_data_descs <- df_data_descs %>% 
#   mutate(description = str_to_sentence(description),
#          variable = str_to_lower(variable),
#          variable = str_squish(variable),
#          variable = str_replace_all(variable, " - ", "-"))
# 
# df <- df %>% mutate(variable = str_to_lower(variable))
# 
# df %>% 
#   anti_join(df_data_descs) %>% distinct(variable)
```


### Plotting

```{r}
df %>% 
  filter(variable == "Population") %>% 
  ggplot(aes(year, value, colour = country)) +
  geom_point() +
  geom_line()
```

What about some of the other complete variables? Start with function to plot these.

Function exported to fns folder

```{r}
# plot_variable <- function(var) {
#   df %>%
#     filter(variable == var,
#            value > 0) %>%
#     ggplot(aes(year, value, colour = country)) +
#     geom_point() +
#     geom_line() +
#     scale_y_continuous(labels = number_format()) +
#     labs(title = paste0("Evolution/adoption of ", var, " over time"),
#          x = NULL,
#          y = NULL)
# }

plot_variable("Mail")
```

What does energy output look like?

```{r}
plot_variable("Energy output")
```

```{r}
plot_variable("Newspapers")
```

Railways - there seems to be a data error for Germany. Imma fix it.

```{r}
plot_variable("Length of Railway line open")
plot_variable("Freight traffic on railways-1")
```

Wow look how Japan speeds ahead in the adoptio of industrial robots over time
```{r}
plot_variable("industrial robots")
```

What about urbanization? Kinda strange - only two points - 1800 and 1850. Same as constraint on executive

```{r}
plot_variable("Urbanization")
plot_variable("Constraint on Executive")

```

### Per capita plot

Maybe I should have a function to plot per person and then plot relative to GDP per capita?

Exported to fns folder

```{r}

# plot_variable_pc <- function(var) {
#   df %>%
#     filter(variable %in% c(var, "Population")) %>%
#     pivot_wider(names_from = variable) %>%
#     rowwise() %>%
#     mutate(
#       value_pc = !!sym(var) / Population,
#       value_pc = replace_na(value_pc, 0)
#     ) %>%
#     filter(value_pc > 0) %>% 
#     ggplot(aes(year, value_pc, colour = country)) +
#     geom_point() +
#     geom_line() +
#     scale_y_continuous(labels = number_format()) +
#     labs(x = NULL,
#          title = paste0("Evolution/adoption of ", 
#                         var, " per capita over time"),
#          y = paste0(var, " per person"))
# }

plot_variable_pc("Energy output")
```

```{r}
plot_variable_pc("Newspapers")
plot_variable("Newspapers")


plot_variable_pc("Phones") +
  scale_y_continuous(labels = percent_format())
```

### gghighlight

```{r}
library(gghighlight)
```

```{r}
country_choices <- c("Sweden", "United Kingdom", "United States", "Spain")

plot_variable_pc("Phones") +
  gghighlight(country %in% country_choices) +
  scale_color_brewer(palette = "Dark2")

plot_variable_pc("Energy output") +
  gghighlight(country %in% country_choices) +
  scale_color_brewer(palette = "Dark2")

plot_variable_pc("Length of Railway line open") +
  gghighlight(country %in% country_choices) +
  scale_color_brewer(palette = "Dark2")
```

What does it mean that Sweden has high literacy rates from 1850, but that primary and secondary enrolment is lower than the UK and the US??

```{r}
plot_variable("Literacy rate") +
  gghighlight(country %in% country_choices) +
  scale_color_brewer(palette = "Dark2") +
  scale_y_continuous(labels = percent_format())

plot_variable("Primary Enrollment") +
  gghighlight(country %in% country_choices) +
  scale_color_brewer(palette = "Dark2") +
  scale_y_continuous(labels = percent_format())

plot_variable("Secondary Enrollment") +
  gghighlight(country %in% country_choices) +
  scale_color_brewer(palette = "Dark2") +
  scale_y_continuous(labels = percent_format())
```

What about the cars?? And transport?

```{r}
plot_variable_pc("Private Cars") +
  gghighlight(country %in% country_choices) +
  scale_color_brewer(palette = "Dark2")


plot_variable_pc("Aviation - PKM") +
  gghighlight(country %in% country_choices) +
  scale_color_brewer(palette = "Dark2")

plot_variable_pc("Aviation-TKM")

plot_variable("Number of Ships-All")
plot_variable("Tonnage-Motorships")

```

Could look at communication technologies?? vs industrial technologies??
```{r}
plot_variable_pc("Mail") +
  gghighlight(country %in% country_choices) +
  scale_color_brewer(palette = "Dark2")

plot_variable_pc("Newspapers") +
  gghighlight(country %in% country_choices) +
  scale_color_brewer(palette = "Dark2")

plot_variable_pc("Phones") +
  gghighlight(country %in% country_choices) +
  scale_color_brewer(palette = "Dark2")

plot_variable_pc("Personal computers") +
  gghighlight(country %in% country_choices) +
  scale_color_brewer(palette = "Dark2")

plot_variable_pc("Televisions") +
  gghighlight(country %in% country_choices) +
  scale_color_brewer(palette = "Dark2")

```

Need to do a GDP plot I think. Gonna try the one with the colour of the dots for time.

```{r}
# df %>% 
#   filter(vaiable %in% c(""))
```

### Plotting trade

```{r}
plot_variable_pc("Trade - Total")

df %>%
  filter(
    variable %in% c("Real GDP", "Imports - 1", "Exports - 1"),
    value > 0
  ) %>%
  mutate(value = case_when(
    variable != "Real GDP" ~ value / 1000,
    TRUE ~ value * 1/100
  )) %>% 
  ggplot(aes(year, value, colour = country)) +
  geom_point() +
  geom_line() +
  facet_wrap(~variable) +
  scale_y_continuous(labels = number_format()) +
  labs(
    # title = paste0("Evolution/adoption of ", var, " over time"),
    x = NULL,
    y = NULL
  )
```

### Trade for sweden

```{r}
df %>% 
  filter(country == "Sweden",
         str_detect(variable, "trade"),
         !str_detect(variable, "sweden"),
         value > 0) %>% 
  mutate(variable = str_remove_all(variable, "trade - "),
         variable = str_to_title(variable),
         time_chunk = year - year %% 3) %>% 
  group_by(time_chunk, variable) %>% 
  mutate(value = mean(value, na.rm = T)) %>% 
  ungroup() %>%
  ggplot(aes(time_chunk, value, colour = variable)) +
  geom_line() +
  geom_point() +
  # scale_y_log10() +
  gghighlight(str_detect(variable, c("Germany", "France", "Denmark", "Spain")))
```

### What about the metrics from AJR and other institutional measures

AJR
- Constraint on Executive
- Protection for capital
- Urbanization

- Type of regime
- Legislative index
- Party legitimacy
- Legislative effectiveness
- Type of head of state
- Legislative effectiveness

```{r}
df %>%
  filter(variable %in% c("Constraint on Executive", "Protection for Capital"),
         !is.na(value)) %>%
  pivot_wider(names_from = variable) %>% 
  filter(year == 1800) %>% 
  ggplot(aes(`Constraint on Executive`, `Protection for Capital`, label = country)) +
  geom_jitter() +
  geom_text()

df %>% 
  filter(variable == "Legislative effectiveness") %>% filter(!is.na(value)) %>% 
  ggplot(aes(value, fill = factor(year))) +
  geom_density()
  
```


```{r}
df %>%
    filter(variable %in% c("Constraint on Executive", "Protection for capital", "Population")) %>%
    pivot_wider(names_from = variable) %>%
    rowwise() %>%
    mutate(across(-c(year, country, Population), .fns = ~ .x / Population)) %>% arrange(desc(year))
```


### What about doing t-SNE or UMAP on the different technologies and then doing a dot plot by country and year

```{r}
df
```

### What about the correlations?

```{r}
df_mod <- df %>% 
  filter(variable %in% c("year",
                         "Real GDP",
                         "Population",
                         "Energy output",
                         "Literacy rate",
                         "Primary Enrollment",
                         "Secondary Enrollment",
                         "Private Cars",
                         "Commercial Cars",
                         "Aviation - PKM",
                         "Aviation-TKM",
                         "Length of Railway line open",
                         "Freight traffic on railways-1",
                         "Passenger traffic on railways-1",
                         "Number of Ships-All",
                         "Tonnage-Motorships",
                         "Mail",
                         "Phones",
                         "Personal Computers",
                         "Mobile Phones",
                         "Telegraph",
                         "Radios",
                         "Newspapers",
                         "Imports - 1",
                         "Exports - 1",
                         "industrial robots",
                         "steel total"))

df_mod <- df_mod %>% 
  filter(!is.na(value)) %>% 
  distinct(.keep_all = T) %>% 
  pivot_wider(names_from = variable, values_from = value, values_fill = 0) %>% 
  janitor::clean_names()
```


### Correlations

First we have the raw correlations - these are simply how the variables are correlated with each other.

What we can see are maybe 3 clusters.

You've got a nice strong correlation between energy output, and transportation like aviation and cars. In that mix are also some communication technologies like Phones and Mail.

Then on the bottom section are railways, newspapers, radios, ships and telegraphs. Similar technologies but from the previous era.

Then obviously literacy, schooling are correlated. Imports and exports. I think it is just a time trend that links imports and secondary education.

Next I want to do a regression controlling for year, grab the residuals, and then do the same correlations.

```{r}
library(tidymodels)

tidy_rec <- recipe(real_gdp ~ ., data = df_mod %>% filter(country == "Sweden")) %>% 
  update_role(c(year, country), new_role = "id") %>% 
  step_log(real_gdp, offset = 1) %>% 
  step_normalize(all_numeric_predictors())
  
df_juiced <- tidy_rec %>% prep() %>% juice()

library(corrr)

corrs <- df_juiced %>% 
  select(-c(year, country)) %>%
  correlate() %>% 
  rearrange()

corrs

set.seed(2021)

corrs %>% 
  network_plot(colours = c("orange", "grey80", "midnightblue"))

```

Could be interesting to do a correlation plot for every decade, and see how it develops over time.


```{r}
df_juiced %>%
  pivot_longer(-c(year, country)) %>% 
  ggplot(aes(year, value, colour = name)) +
  geom_point() +
  geom_line()
```

### Recipe for controlling for a time trend

Splits:

```{r}
split <- initial_split(df_mod, strata = year)

set.seed(2021)
df_train <- training(split)
df_test <- testing(split)
df_folds <- vfold_cv(df_train)
```


```{r}
lasso_rec <- recipe(real_gdp ~ ., data = df_mod) %>% 
  update_role(c(country), new_role = "id") %>% 
  step_log(real_gdp, offset = 1) %>% 
  step_normalize(all_numeric_predictors(), -year)

lasso_juiced <- lasso_rec %>% prep() %>% juice()

lasso_spec <- linear_reg(penalty = tune(), mixture = 1) %>% 
  set_mode("regression") %>% 
  set_engine("glmnet") 

lasso_workflow <- 
  workflow() %>% 
  add_recipe(lasso_rec) %>% 
  add_model(lasso_spec) 

lasso_grid <- tidyr::crossing(penalty = 10^seq(-6, -1, length.out = 20)) 

lasso_tune <- 
  tune_grid(lasso_workflow, resamples = df_folds, grid = lasso_grid) 

lasso_tune %>% autoplot()

lasso_best <- lasso_tune %>% select_best(., "rsq")

lasso_final_wf <- finalize_workflow(lasso_workflow, lasso_best)

lasso_final_fit <- last_fit(lasso_final_wf, split = split)
```

What does it look like?

```{r}
extract_fit_parsnip(lasso_final_fit) %>% tidy() %>% 
  filter(term != "(Intercept)") %>% 
  slice_max(abs(estimate), n = 20) %>% 
  mutate(term = fct_reorder(term, estimate)) %>% 
  ggplot(aes(estimate, term, fill = estimate > 0)) +
  geom_col()

lm(real_gdp ~ ., data = lasso_juiced %>% select(!country)) %>% 
  tidy() %>% 
  filter(term != "(Intercept)") %>% 
  slice_max(abs(estimate), n = 20) %>% 
  mutate(term = fct_reorder(term, estimate)) %>% 
  ggplot(aes(estimate, term, fill = estimate > 0)) +
  geom_col()

```

### Energy adoption over time

```{r}
df %>% 
  filter(between(year, 1900, 1950),
         variable %in% c("Energy output", "Population")) %>% 
  pivot_wider(names_from = variable) %>% 
  rowwise() %>% 
  mutate(value_pc = `Energy output` / Population,
         value_pc = replace_na(value_pc, 0)) %>% 
  ggplot(aes(year, value_pc, colour = country)) +
  geom_point() + 
  geom_line() +
  gghighlight::gghighlight(country == "Sweden") +
  labs(x = NULL,
       y = "Energy output in 10mW per capita",
       title = "Sweden has a relatively high energy output per capita")
```

